Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-10 paper/disc20clustering.tex
wip
Improvement of lem:1req, now regular proof instead of by contradiction

\noindent
\textbf{Component-based algorithm.}
Now we describe the algorithm \TAlg.
The algorithm partitions nodes into components, and
initially, each node is isolated (belongs to its own component).
For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
If no such partitioning exists, \TAlg resets all components to trivial components, resets all counters to $0$, and ends the phase.

@@

\begin{proof}
	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.

@@

	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a component of size $3$ participates in a merge.
	After the merge, we would have a component of size at least $4$, and in this case, no component respecting partitioning exists, a contradiction.

	It remains to consider the case where both $U$ and $V$ are of type $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists. 
	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists. In this case, \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes~of~$W$.
\end{proof}

>>>>

\noindent
\textbf{Component-based algorithm.}
We introduce the algorithm \TAlg that we analyze in this section.
The algorithm partitions nodes into components, and
initially, each node is isolated (belongs to its own component).
For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
If no such partitioning exists, \TAlg resets all components to trivial components, resets all counters to $0$, and ends the phase.

@@

\begin{proof}
	If no component respecting partition exists after the merge of components, \TAlg simply ends the phase and performs no reconfiguration.
	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.

@@

	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
	If either $U$ or $V$ is of type $C_3$, a component of size $3$ participates in a merge, and we have a component of size at least $4$, and \TAlg ends the phase with no reconfiguration.

	It remains to consider the case where both $U$ and $V$ are of type $C_2$.
	If $(u,v)$ both belong to components of size $2$, then the merged component has size $4$, and \TAlg incurs no cost. 
	Otherwise, if one of $u,v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.

	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
	Note that if $C_1$-type cluster does not exist, then no component respecting partitioning exists.
	Otherwise, \TAlg performs one swap --- it exchanges the nodes $u$ and $v$ with any two nodes~of~any cluster of type $C_1$.

	In each case, we showed that a partition is reachable in at most two swaps.
\end{proof}
